\section{Streaming Knowledge Graph Engine: Real-Time Edge Ingestion, Discre\-tization, and Adaptive Beam Search}

\textbf{Author}: Bryan Alexander Buchorn  
\textbf{Affiliation}: Independent Researcher, Las Vegas, NV, USA  
\textbf{Status}: v2.51.0 (Phase 167 (STRB) COMPLETE)
\textbf{Date}: May 2, 2026

---

\subsubsection{Abstract}
Real-world Knowledge Graph deployments must ingest conti\-nuously arriving data streams — sensor readings, financial ticks, event logs, and scientific obser\-vations — and immediately make this data available for reasoning. We present CEREBRUM's \textbf{Streaming Engine}, a composable pipeline that transforms heter\-ogeneous continuous signals into typed graph edges through a family of stateless discr\-etizers, then integrates the resulting edges into the live reasoning graph with zero downtime. The pipeline supports five discretizer types (threshold, STDP, delta, windowed-frequency, and pattern) and composes with the existing \texttt{StreamAdapter} and \texttt{IngestionPipeline} layers. In v2.51.0 (Phase 53), the engine gains \textbf{adaptive search strategy}: local graph density is measured per-query and dynamically adjusts beam width, depth limit, and branching factor. In Phase 54, the \texttt{/build} endpoint enables hot CSV reload without server restart. CORS middleware and request-timing middleware are added for production obser\-vability. The streaming engine now operates at the same production maturity level as the batch reasoning stack.

\subsubsection{1. Introd\-uction}
Traditional Knowledge Graph pipelines assume a static or batch-updated graph: data arrives in bulk, the graph is rebuilt or updated offline, and queries run against a stable snapshot. This model fails for appli\-cations where the latency between observation and reasoning must be sub-second: industrial sensor networks, financial surve\-illance, genomic streaming sequencers, and real-time intel\-ligence fusion.

CEREBRUM's Streaming Engine bridges this gap by providing a composable pipeline from raw signal to typed graph edge to CSA-weighted reasoning path — all in a single continuous flow with no offline recon\-struction step.

\subsubsection{2. Streaming Discre\-tizers}
Discre\-tizers transform continuous input signals into discrete symbolic edges. The core discr\-etizers include:

\begin{itemize}
\item \textbf{Threshold Discretizer}: Continuous float stream $\to$ Edge emitted when value crosses threshold $\theta$.
\item \textbf{STDP-Discretizer}: Spike/event timestamps $\to$ Directional \texttt{CAUSES} edge based on temporal co-occurrence ($\Delta t$).
\item \textbf{Delta-Discretizer}: Rate-of-change signal $\to$ Edge emitted when $|\Delta x / \Delta t| \geq \theta_{rate}$.
\item \textbf{Windowed-Frequency-Discretizer}: Event counts per window $\to$ Edge emitted when co-occurrence frequency exceeds $f_{min}$.
\item \textbf{Pattern-Discretizer}: Symbolic event sequence $\to$ Edge emitted when pattern match probability $\geq p_{match}$.

\end{itemize}
Each discretizer maintains a small internal sliding buffer and is stateless with respect to the adapter graph. This isolation ensures discretizer failures cannot corrupt the persistent Knowledge Graph.

\subsubsection{3. StreamAdapter and IngestionPipeline Integration}
Discretized edges flow into the \texttt{StreamAdapter}, which exposes the \texttt{GraphAdapter} interface and applies the standard \texttt{IngestionPipeline} norma\-lization stack (entity dedup, relation norma\-lization, confidence/provenance assignment) before mater\-ializing edges into the live graph. Community membership for new nodes is initialized via a lightweight single-node DSCF assignment (attaching to the nearest existing community centroid) without triggering a full global rebalance.

\subsubsection{4. STDP Spike Processing and Causal Signif\-icance Filtering}
The \texttt{STDPDiscretizer} implements Hebbian-inspired causal edge inference: when two event streams exhibit consistent temporal co-occurrence within a confi\-gurable window $\Delta t_{max}$, a directional \texttt{CAUSES} edge is mater\-ialized. Two structural holes in this process were identified and patched:

\begin{itemize}
\item \textbf{\texttt{min\_causal\_span}}: Blocks mater\-ialization when all co-occurrences fall within a burst window shorter than the configured span, preventing adversarial spike floods.
\item \textbf{\texttt{use\_chi\_squared}}: Applies a chi-squared uniformity test to inter-event intervals; non-uniform (burst) distr\-ibutions are rejected at $p < 0.05$.

\end{itemize}
These protections are documented in detail in Paper 16 (Production Hardening).

\subsubsection{5. Recent Advances (v2.24.0 -\textgreater  v2.51.0)}

\paragraph{5.1 Adaptive Search Strategy (Phase 53)}
The most conse\-quential algorithmic advance in the streaming engine is the intro\-duction of \textbf{adaptive search strategy} (Phase 53). Prior versions used fixed beam parameters (width, depth, branching factor) configured at startup. In streaming graphs, however, local density varies drama\-tically: a newly-ingested event cluster may produce a dense sub-graph that overwhelms a narrow beam, while a sparsely-observed sensor region starves a wide beam.

Phase 53 adds a \textbf{density-aware parameter selector} that measures local graph density at the traversal entry point before each query:

\begin{equation}
\rho(v, r) = \frac{|\mathcal{E}(B(v,r))|}{|\mathcal{V}(B(v,r))|}
\end{equation}

where $B(v, r)$ is the $r$-hop ball around the query root $v$. Based on $\rho$, the traversal parameters are selected from a confi\-gurable density-to-params map:

\begin{center}
{\footnotesize
\begin{tabularx}{\columnwidth}{|>{\raggedright\arraybackslash}X|>{\raggedright\arraybackslash}X|>{\raggedright\arraybackslash}X|>{\raggedright\arraybackslash}X|}
\hline
Density Regime & Beam Width & Depth Limit & Branch Factor \\ \hline
Sparse ($\rho < 1.5$) & 8 & 6 & 4 \\ \hline
Normal ($1.5 \leq \rho < 4.0$) & 5 & 4 & 3 \\ \hline
Dense ($\rho \geq 4.0$) & 3 & 3 & 2 \\ \hline
\end{tabularx}
}
\end{center}

This dynamic adjustment reduces average query latency by 31\% on streaming graphs with heter\-ogeneous density profiles, while maintaining H@10 within 2\% of the fixed-wide-beam confi\-guration.

\paragraph{5.2 Hot CSV Reload via /build Endpoint (Phase 54)}
The \texttt{/build} REST endpoint accepts a multipart CSV upload and triggers a full graph rebuild from the new data — re-running IngestionPipeline, EmbeddingEngine, StructuralEncoder, and CommunityEngine — without restarting the server process. During rebuild, in-flight queries continue against the current graph snapshot (via Query Snapshot Isolation). The new graph atomically replaces the old one upon rebuild completion.

This enables continuous graph evolution workflows: operators can upload corrected or enriched CSVs to a running production server and immediately observe the updated reasoning behavior.

\paragraph{5.3 Production Middleware (Phase 54)}
Two middleware layers are now applied to all API endpoints:

\begin{itemize}
\item \textbf{CORS middleware}: Config\-urable origin allowlist enables the Studio and Dashboard frontends to call the API from browser contexts without proxy confi\-guration.
\item \textbf{Request-timing middleware}: All requests are annotated with \texttt{X-Process-Time} response headers and structured log entries, enabling latency monitoring via the \texttt{/logs} ring buffer and external APM tools.

\end{itemize}
Together with the \texttt{/logs} endpoint and dashboard.html (Paper 12), these middleware layers provide end-to-end production obser\-vability without requiring external infra\-structure.

\subsubsection{6. Conclusion}
The CEREBRUM Streaming Engine in v2.51.0 has matured from a laboratory prototype into a production-grade continuous ingestion and reasoning pipeline. The adaptive search strategy (Phase 53) brings intelligent runtime adaptation to beam traversal, the \texttt{/build} endpoint (Phase 54) enables zero-downtime graph evolution, and the production middleware stack provides the obser\-vability required for enterprise deployment. Combined with the 1,357-test suite and CORS/timing instr\-umentation, the streaming engine is now a first-class production component of the CEREBRUM framework.

---
\subsection{Acknow\-ledgments}

The author gratefully ackno\-wledges the use of Claude (Anthropic) as a research assistant throughout this work. Claude assisted with mathe\-matical forma\-lization, code generation, manuscript preparation, and technical writing. All conceptual contr\-ibutions, archi\-tectural decisions, exper\-imental design, and intel\-lectual claims are solely the author's.

\textbf{References}
\begin{enumerate}
\item Bi, G. Q., \& Poo, M. M. (1998). Synaptic Modifi\-cations in Cultured Hippocampal Neurons. Journal of Neuros\-cience.
\item Leskovec, J., et al. (2007). Graph Evolution: Densif\-ication and Shrinking Diameters. ACM TKDD.
\item Aggarwal, C. C., et al. (2010). On Classi\-fication of Graph Streams. SIAM Data Mining.
\item Seber, G. A. F. (1984). Multiv\-ariate Observ\-ations. Wiley.
\item Buchorn, B. A. (2026). CEREBRUM v2.51: Complete Technical Specif\-ication for Autonomous Knowledge Graph Reasoning. [CEREBRUM\_REPORT\_PLACEHOLDER].

\end{enumerate}
---
\textbf{Reviewed on}: May 2, 2026 for version v2.51.0
