\section{CSA: Community-Structured Attention for Knowledge Graph Reasoning}

\textbf{Author}: Bryan Alexander Buchorn  
\textbf{Affiliation}: Independent Researcher, Las Vegas, NV, USA  
\textbf{Status}: v2.51.0 (Phase 167 (STRB) COMPLETE)
\textbf{Date}: May 2, 2026

---

\subsubsection{Abstract}
We propose \textbf{Community-Structured Attention (CSA)}, an attention mechanism that enables multi-hop reasoning over large Knowledge Graphs (KGs) without the $O(N^2)$ complexity of global attention matrices. CSA maps the structural components of the Transformer archi\-tecture \cite{vaswani2017attention} directly onto graph operations, utilizing community partitions as discrete "Attention Heads." We define a unified scoring function that integrates semantic similarity, community-level topology, and structural centrality. Benchmark results on the \textbf{Hetionet} \cite{hetionet2017} biomedical dataset demonstrate that CSA achieves a Mean Reciprocal Rank (MRR) of \textbf{0.68}, a \textbf{+183\% improvement} over breadth-first search baselines. Furthermore, on the \textbf{MetaQA 3-hop} \cite{metaqa2017} reasoning task, CSA improves MRR by \textbf{+350\%}, demon\-strating superior beam steering in deep multi-hop traversals while maintaining full "Glass-Box" inter\-pretability. As of v2.51.0, the CSA formula has been expanded to 10 parameters covering temporal decay, node recency, synthesis-density penalty, and grounding confidence, with both batch (CSAParameterLearner) and online per-community (MetaParameterLearner) learning, achieving MetaQA canonical results of H@1=46.1\%/30.0\%/12.5\% across 1-, 2-, and 3-hop tasks.

\subsubsection{1. Introd\-uction}
The dominance of Transformer archi\-tectures in Natural Language Processing has inspired attempts to apply similar attention-based principles to graph structures. However, Graph Attention Networks (GATs) \cite{velickovic2018gat} typically operate on local ego-networks and struggle with global structural context. CSA addresses this by introducing a "Soft Community Constraint," where attention weights are influenced by the membership of nodes in pre-computed structural partitions (DSCF/TSC).

\subsubsection{2. The Cerebrum Mapping}
CSA is built on a direct functional analogy to the Transformer \cite{vaswani2017attention}:
\begin{itemize}
\item \textbf{Communities} act as \textbf{Attention Heads}, focusing the search on specific semantic neigh\-borhoods.
\item \textbf{Centrality Features} (PageRank, Betweenness) serve as \textbf{Positional Encodings}, providing structural context.
\item \textbf{Traversal Paths} function as a \textbf{KV Cache}, memoizing the reasoning history.

\end{itemize}
\subsubsection{3. Methodology}

\paragraph{3.1 The CSA Formula}
The attention weight $a(u,v,k)$ for an edge from $u$ to $v$ at hop $k$ is defined as:
$$\begin{aligned}
a(u,v,k) = \sigma( \& \alpha \cdot \mathcal{S}\_{sem}(u,v) + \beta \cdot \mathcal{S}\_{com}(u,v) + \\
\& \gamma \cdot w\_{rel} - \delta \cdot d\_{norm}(u,v) + \epsilon \cdot \phi(k) )
\end{aligned}$$

\paragraph{3.2 The Community Signal ($\mathcal{S}_{com}$)}
Unlike GATs \cite{velickovic2018gat} which treat all neighbors equally, CSA scales weights based on community topology:
\begin{itemize}
\item \textbf{Intra-community}: $1.0$
\item \textbf{Adjacent-community}: $0.5$
\item \textbf{Distant-community}: $e^{-\lambda d_{com}}$

\end{itemize}
\subsubsection{4. Enterprise Hardening (v2.51.0)}
The v2.51.0 release introduces \textbf{Adaptive Parameter Learning}, utilizing a \textbf{MetaParameterLearner} to auton\-omously adjust the $(\alpha, \beta, \gamma, \delta, \epsilon)$ coeff\-icients per-community based on query feedback. This closes the gap between zero-shot and supervised performance without the need for global retraining.

\subsubsection{5. Conclusion}
CSA provides a scalable, Interp\-retable AI (XAI) alternative to black-box graph embeddings. By grounding attention in the structural consensus of the graph, it enables complex multi-hop reasoning that is both compu\-tationally efficient and mathe\-matically verifiable. In CEREBRUM v2.51.0, the 10-parameter CSA formula with online per-community learning achieves MetaQA H@1 of 46.1\% (1-hop), 30.0\% (2-hop), and 12.5\% (3-hop), alongside WebQSP H@1=6.27\%, H@10=20.84\%, and MRR=10.66\% — estab\-lishing CSA as a competitive and inter\-pretable alternative to embedding-based KG reasoning.

---

\subsection{6. Recent Advances (v2.24.0 -\textgreater  v2.51.0)}

The CSA formula and its associated learning infra\-structure have undergone significant expansion since v2.24.0. The following describes the key advances relevant to this paper.

\textbf{10-Parameter CSA Formula (Phase 43/45).} The original 5-parameter formula \texttt{(alpha, beta, gamma, delta, epsilon)} covering semantic similarity, community score, edge-type weight, distance penalty, and hop decay has been extended to 10 parameters:

\begin{lstlisting}
a(u,v,k) = sigmoid(
    alpha   * sim          # semantic similarity (cosine)
  + beta    * cs           # community score (structural membership)
  + gamma   * etw          # edge-type weight
  - delta   * nd           # normalised distance penalty
  + epsilon * hd           # hop decay
  + zeta    * pr_v         # PageRank prior
  + eta     * td           # temporal decay
  + iota    * nr_v         # node recency
  - mu      * sd           # synthesis-density penalty
  + theta   * grounding    # confidence / grounding score
)
\end{lstlisting}

Default weights: \texttt{(0.4, 0.4, 0.1, 0.05, 0.05, 0.1, 0.1, 0.05, 0.1, 1.0)}. The synthesis-density penalty (\texttt{-mu*sd}) is parti\-cularly significant: it prevents over-reliance on REM-synthesized Synaptic Bridge edges, maintaining reasoning trans\-parency.

\textbf{Unified ReasoningLogit Vector.} All 10 features are bundled into a \texttt{ReasoningLogit} dataclass that threads through scoring, learning, and feedback logging. This ensures that the full feature vector is available for inspection at every hop, preserving the "Glass-Box" property of the original design.

\textbf{Batch and Online Parameter Learning (Phase 22/45/48).} Two learning regimes are now fully implemented:
\begin{itemize}
\item \texttt{CSAParameterLearner.fit()}: batch gradient descent over accumulated (positive, negative) path pairs, triggered via \texttt{POST /retrain}. Updates the global 10-parameter prior.
\item \texttt{MetaParameterLearner}: online SGD per community, updated on every \texttt{POST /feedback} call. Enables community-specific adaptation without global retraining.

\end{itemize}
\textbf{Params Persistence (Phase 47).} \texttt{MetaParameterLearner.to\_dict()} / \texttt{from\_dict()} enables checkpoint and restore via \texttt{POST /params}. The \texttt{--params-file} CLI flag loads a checkpoint at server startup, enabling zero-downtime parameter rollback.

\textbf{Benchmark Results.}

\begin{center}
{\footnotesize
\begin{tabularx}{\columnwidth}{|>{\raggedright\arraybackslash}X|>{\raggedright\arraybackslash}X|>{\raggedright\arraybackslash}X|}
\hline
Dataset & Metric & v2.51.0 \\ \hline
MetaQA 1-hop & H@1 / H@10 & 46.1\% / 96.6\% \\ \hline
MetaQA 2-hop & H@1 / H@10 & 30.0\% / 86.3\% \\ \hline
MetaQA 3-hop & H@1 / H@10 & 12.5\% / 50.3\% \\ \hline
WebQSP OPT & H@1 / H@10 / MRR & 6.27\% / 20.84\% / 10.66\% \\ \hline
\end{tabularx}
}
\end{center}

\subsection{7. Phase 55 Advances}

\textbf{GraphSAGE Neighb\-ourhood Smoothing (Phase 55).} \texttt{smooth\_with\_graphsage(embeddings, G)} applies a one-pass mean neigh\-bourhood aggregation after base encoding:

\begin{equation}
\tilde{\mathbf{e}}_v = \frac{1}{1+|\mathcal{N}(v)|}\left(\mathbf{e}_v + \sum_{u \in \mathcal{N}(v)} \mathbf{e}_u\right)
\end{equation}

The enriched embeddings make the \texttt{alpha} (semantic similarity) term in the CSA formula signi\-ficantly more discr\-iminating — nodes in the same community share more similar neigh\-bourhood-aggregated repre\-sentations. \texttt{CerebrumGraph.build(use\_graphsage=True)} enables smoothing autom\-atically after base encoding. Complexity is $O(|E| \times d)$ where $d$ is the embedding dimension.

\textbf{TemporalCalibrator (Phase 55).} Grid-searches \texttt{eta} (temporal decay) and \texttt{iota} (node recency) against a labelled validation set to maximise Recall@K. The \texttt{calibrate()} method enumerates a parameter grid, calls \texttt{measure\_recall()} at each point, and applies the best-found parameters to the CSAEngine. A \texttt{try/finally} block guarantees original parameters are restored if calibration is interrupted — ensuring that a failed calibration run never leaves the CSAEngine in a partially-modified state.

\textbf{Engram-Steered Traversal (Phase 55).} \texttt{Engram} tracks relation-sequence patterns from previous successful Engram traces. \texttt{EngramTraversal.\_prune\_candidates()} applies:

\begin{equation}
s_\text{eff}(c) = s(c) \times (1 + \lambda_\text{engram} \cdot \text{affinity}(\text{rel\_seq}))
\end{equation}

where \texttt{affinity} is derived from accumulated \texttt{\_counts}. This biases beam search toward known-productive reasoning chains without modifying graph structure. The cache is durable — \texttt{save(path)} serializes to JSON and \texttt{load(path)} restores counts on restart, so learned relation patterns survive process restarts.

---
\subsection{Acknow\-ledgments}

The author gratefully ackno\-wledges the use of Claude (Anthropic) as a research assistant throughout this work. Claude assisted with mathe\-matical forma\-lization, code generation, manuscript preparation, and technical writing. All conceptual contr\-ibutions, archi\-tectural decisions, exper\-imental design, and intel\-lectual claims are solely the author's.

\textbf{References}
\begin{enumerate}
\item Veličković, P., et al. (2018). Graph Attention Networks. ICLR.
\item Vaswani, A., et al. (2017). Attention is All You Need. NIPS.
\item Reimers, N., \& Gurevych, I. (2019). Sentence-BERT: Sentence Embeddings using Siamese BERT-Networks. EMNLP.
\item Bordes, A., et al. (2013). Translating embeddings for modeling multi-relational data. NIPS.
\item Sun, Z., et al. (2019). RotatE: Knowledge Graph Embedding by Relational Rotation in Complex Space. ICLR.
\item Edge, D., et al. (2024). From Local Retrieval to Global Explanation of Text Graphs. Microsoft Research.
\item Himmelstein, C. S., et al. (2017). Systematic integration of biomedical knowledge prioritizes drugs for infla\-mmation. eLife.
\item Zhang, Y., et al. (2018). Variational Reasoning for Question Answering over Knowledge Graphs. ICLR.
\item Wang, Q., et al. (2017). Knowledge Graph Embedding: A Survey of Approaches and Applic\-ations. IEEE TKDE.
\item Buchorn, B. A. (2026). CEREBRUM v2.51: Complete Technical Specif\-ication for Autonomous Knowledge Graph Reasoning. [CEREBRUM\_REPORT\_PLACEHOLDER].
\item Hamilton, W., Ying, Z., \& Leskovec, J. (2017). Inductive Repres\-entation Learning on Large Graphs. NeurIPS.

\end{enumerate}
---
\textbf{Reviewed on}: May 2, 2026 for version v2.51.0
