\section{PAPER 032: GraphAdapter remove\_edge Protocol — A Formal Edge-Removal Contract for Knowledge Graph Adapters}

\textbf{Author}: Bryan Alexander Buchorn  
\textbf{Affiliation}: Independent Researcher, Las Vegas, NV, USA  
\textbf{Status}: v2.52.0 (Phase 172 (STRB) COMPLETE)
\textbf{Date}: May 2, 2026

\textbf{CEREBRUM Phase 80}

---

\subsection{Abstract}

We present the \textbf{GraphAdapter \texttt{remove\_edge()} Protocol}, a Phase 80 change that promotes edge removal from an ad-hoc capability to a formally defined method in the \texttt{GraphAdapter} base class. Prior to Phase 80, \texttt{ProvenanceLedger.rollback\_batch()} checked for edge-removal capability via \texttt{hasattr(adapter, "remove\_edge")} and silently skipped rollback if the method was absent. This silent-skip pattern allowed custom adapter subclasses to silently fail rollback without any error signal. Phase 80 removes the guard and defines \texttt{remove\_edge(u, v, relation)} as a protocol method on \texttt{GraphAdapter} — non-abstract, raising \texttt{NotImplementedError} by default, overridden in all six built-in adapters. Custom adapters must explicitly override the method or accept that rollback will raise \texttt{NotImplementedError}, making the failure visible. This is a breaking change for custom adapters that previously relied on the silent-skip behavior.

---

\subsection{1. Motivation: Silent Failure in Rollback}

The ProvenanceLedger (Phase 76) enables targeted rollback of mater\-ialized edges. Rollback is only useful if the underlying adapter can actually remove edges. The original imple\-mentation used a defensive \texttt{hasattr()} check:

\begin{lstlisting}
# Pre-Phase 80 (fragile)
if hasattr(adapter, "remove_edge"):
    adapter.remove_edge(u, v, relation)
else:
    logger.warning("Adapter does not support remove_edge; skipping rollback")
\end{lstlisting}

This pattern has two failure modes:
\begin{enumerate}
\item \textbf{Silent no-op}: A custom adapter imple\-menting a different interface (e.g., \texttt{delete\_edge}) would fail the \texttt{hasattr} check, silently skip rollback, and report success.
\item \textbf{Untestable absence}: Tests that mock \texttt{adapter} without adding \texttt{remove\_edge} would pass, masking the missing imple\-mentation.

\end{enumerate}
---

\subsection{2. Protocol Definition}

The Phase 80 change to \texttt{core/graph\_adapter.py}:

\begin{lstlisting}
class GraphAdapter(ABC):
    # ... existing abstract methods ...

    def remove_edge(self, u: str, v: str, relation: str) -> None:
        """Remove edge (u, v, relation) from the graph.

        All built-in adapters override this method.
        Custom subclasses must override it to enable ProvenanceLedger rollback.
        """
        raise NotImplementedError(
            f"{type(self).__name__} does not implement remove_edge(). "
            "Override this method to enable ProvenanceLedger rollback."
        )
\end{lstlisting}

The method is \textbf{non-abstract} (does not use \texttt{@abstractmethod}) to avoid breaking existing custom subclasses on import. The \texttt{NotImplementedError} is raised at call time, not at subclass definition time.

---

\subsection{3. Built-in Adapter Implem\-entations}

All six built-in adapters implement \texttt{remove\_edge}:

\begin{center}
\begin{tabularx}{\columnwidth}{|>{\raggedright\arraybackslash}X|>{\raggedright\arraybackslash}X|}
\hline
Adapter & Implem\-entation \\ \hline
\texttt{NetworkXAdapter} & \texttt{G.remove\_edge(u, v)} with edge matching by relation key \\ \hline
\texttt{Neo4jAdapter} & Cypher \texttt{MATCH (u)-[r:RELATION]-\textgreater (v) DELETE r} \\ \hline
\texttt{RDFAdapter} & \texttt{graph.remove((subject, predicate, obj))} \\ \hline
\texttt{CSVAdapter} & In-memory edge list filter + optional file flush \\ \hline
\texttt{StreamAdapter} & Delegates to wrapped \texttt{base\_adapter.remove\_edge()} \\ \hline
\texttt{RemoteCerebrumAdapter} & \texttt{DELETE /edges} REST call with signature \\ \hline
\end{tabularx}
\end{center}

---

\subsection{4. Migration Guide for Custom Adapters}

Any custom \texttt{GraphAdapter} subclass must now explicitly handle edge removal:

\begin{lstlisting}
# Option A: Implement it
class MyAdapter(GraphAdapter):
    def remove_edge(self, u: str, v: str, relation: str) -> None:
        self._edges = [(a, b, r) for (a, b, r) in self._edges
                       if not (a == u and b == v and r == relation)]

# Option B: Declare it unsupported
class MyAdapter(GraphAdapter):
    def remove_edge(self, u: str, v: str, relation: str) -> None:
        raise NotImplementedError("MyAdapter does not support edge removal.")
\end{lstlisting}

Option B is equivalent to the previous silent-skip behavior, but now raises explicitly rather than silently doing nothing.

---

\subsection{5. ProvenanceLedger Changes}

\begin{lstlisting}
# Post-Phase 80 (no hasattr guard)
def rollback_batch(self, batch_id: str, adapter: GraphAdapter) -> int:
    batch = self._batches[batch_id]
    for edge in batch.edges:
        adapter.remove_edge(edge.u, edge.v, edge.relation)  # raises if not implemented
    batch.rolled_back = True
    return len(batch.edges)
\end{lstlisting}

If \texttt{adapter.remove\_edge()} raises, the exception propagates to the caller. \texttt{ProvenanceLedger} does not catch it — callers are expected to handle \texttt{NotImplementedError} if they choose to continue on partial rollback.

---
\textbf{Copyright © 2026 Bryan Alexander Buchorn. All Rights Reserved.}

---
\subsection{Acknow\-ledgments}

The author gratefully ackno\-wledges the use of Claude (Anthropic) as a research assistant throughout this work. Claude assisted with mathe\-matical forma\-lization, code generation, manuscript preparation, and technical writing. All conceptual contr\-ibutions, archi\-tectural decisions, exper\-imental design, and intel\-lectual claims are solely the author's.

\textbf{Reviewed on}: May 2, 2026 for version v2.52.0
