
\section*{Abstract}

We present \textbf{AutoApprover}, a three-tier automated decision engine for \texttt{ResearchFinding} objects produced by CEREBRUM's \texttt{ResearchAgent}. AutoApprover replaces manual finding review in the Autonomous Discovery Loop (Phase 74) with a principled decision stack: (1) \textbf{hard gates} that reject structurally invalid findings instantly; (2) an \textbf{online logistic SGD classifier} operating on a 16-dimensional feature vector; (3) an optional \textbf{LLM semantic fallback} for borderline cases. The 16-feature vector incorporates confidence, discovery potential, community topology, TriangulationEngine scores (Phase 72), and novelty metrics. Online \texttt{fit()} from confirmed decisions enables continuous improvement without retraining cycles. AutoApprover checkpoints via \texttt{to\_dict()} / \texttt{from\_dict()}, enabling warm restart after pod restarts. In the Autonomous Discovery Loop, AutoApprover maintains a rolling approval rate that drives the circuit breaker, ensuring graph quality degrades gracefully under noise.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{1. Motivation: The Manual Approval Bottleneck}

\texttt{ResearchAgent} (Phase 51) discovers candidate missing links in the Knowledge Graph by running \texttt{HypothesisEngine} across underrepresented community boundaries. In the original design, all findings were queued for human review. This creates a throughput bottleneck that prevents fully autonomous operation.

A naive threshold on the raw \texttt{confidence} field is insufficient: high-confidence findings may duplicate existing edges, contradict strong existing evidence, or belong to over-saturated communities where further materialization adds noise rather than information. A decision engine must consider multiple dimensions simultaneously.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{2. Three-Tier Decision Stack}

\subsection*{2.1 Tier 1: Hard Gates}

Hard gates fire before any ML computation. A finding is immediately \textbf{REJECTED} if:

\begin{table}[h]
\small\centering
\begin{tabular}{ll}
\toprule
Gate & Condition \\
\midrule
Literature block & \texttt{finding.literature\_status} in \texttt{\{"BLOCKED", "RETRACTED"\}} \\
Missing validation & \texttt{finding.validation\_report} is \texttt{None} \\
Contradiction threshold & \texttt{finding.metadata["contradiction\_score"] > HARD\_THRESHOLD} (default 0.9) \\
\bottomrule
\end{tabular}
\end{table}


Hard gate rejections are cheap (O(1)) and prevent obviously-bad findings from consuming classifier compute.

\subsection*{2.2 Tier 2: Online SGD Classifier}

A logistic regression classifier operating on a 16-dimensional feature vector:

\begin{table}[h]
\small\centering
\begin{tabular}{lll}
\toprule
# & Feature & Source \\
\midrule
1 & \texttt{confidence} & HypothesisEngine output \\
2 & \texttt{discovery\_potential} & DiscoveryCalibrator weight $\times$ raw potential \\
3 & \texttt{gap\_score} & Community structural gap \\
4 & \texttt{community\_distance} & Hop distance between source/target communities \\
5 & \texttt{local\_density} & Edge density around proposal \\
6 & \texttt{lit\_status\_ordinal} & literature_status encoded as int \\
7 & \texttt{novelty\_score} & 1 - similarity to existing graph edges \\
8 & \texttt{engram\_affinity} & Engram pattern match strength \\
9 & \texttt{path\_count} & Number of independent supporting paths \\
10 & \texttt{contradiction\_score} & ContradictionResolver net_evidence_score \\
11--14 & \texttt{triangulation\_*} & TriangulationEngine P1--P4 scores (Phase 72) \\
15 & \texttt{seeded\_by\_research} & Boolean: finding originated from ResearchAgent scan \\
16 & \texttt{seeded\_by\_external} & Boolean: finding triggered by external literature signal \\
\bottomrule
\end{tabular}
\end{table}


SGD update rule on confirmed decisions:
\begin{lstlisting}[language=Python]
approver.fit(finding, label=True)   # confirmed approval
approver.fit(finding, label=False)  # confirmed rejection
\end{lstlisting}

\subsection*{2.3 Tier 3: LLM Semantic Fallback}

When the classifier score is within \texttt{[threshold - margin, threshold + margin]} (borderline), an optional LLM call is made via \texttt{LLMFallback.evaluate(finding)}. The LLM receives a structured prompt with the entity pair, relation, confidence, and top supporting paths, and returns APPROVE / REJECT / REVIEW.

LLM fallback is disabled by default and requires explicit wiring:
\begin{lstlisting}[language=Python]
approver = AutoApprover(llm_fallback=AnthropicFallback(model="claude-sonnet-4-6"))
\end{lstlisting}

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{3. Checkpoint and Restore}

\begin{lstlisting}[language=Python]
state = approver.to_dict()    # JSON-serializable checkpoint
approver2 = AutoApprover.from_dict(state)   # restore on restart
\end{lstlisting}

The checkpoint persists the SGD weight vector, threshold, and decision history count.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{4. REST API}

\begin{lstlisting}
GET  /research/auto-approver          → current weights, threshold, decision counts
POST /research/auto-approver          → partial update (threshold, fallback config)
\end{lstlisting}

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{5. Integration with Autonomous Discovery Loop}

\texttt{AutonomousDiscoveryLoop} passes each finding through \texttt{approver.decide(finding)}:
\begin{itemize}
  \item \textbf{APPROVE} $\to$ \texttt{research\_agent.approve(finding)} $\to$ edges materialized
  \item \textbf{REJECT} $\to$ \texttt{research\_agent.reject(finding)} $\to$ finding discarded
  \item \textbf{REVIEW} $\to$ added to review queue (no immediate action)
\end{itemize}


The rolling approval rate (APPROVE / (APPROVE + REJECT)) feeds the circuit breaker.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{6. References}

\begin{itemize}
  \item [bottou2010large] Bottou, L. (2010). Large-scale machine learning with stochastic gradient descent. \textit{COMPSTAT}, 177--186.
\end{itemize}
