
\section*{Abstract}

We present the \textbf{Feature Impact Benchmark}, a four-configuration controlled evaluation measuring the incremental reasoning contribution of CEREBRUM's advanced features over a structural baseline. The benchmark evaluates Hits@1, Hits@5, and MRR on the canonical toy_graph.csv fixture (21 nodes, 30 edges) across four progressively richer configurations: (1) \textbf{baseline} --- pure CSA + BeamTraversal; (2) \textbf{+engram} --- adds Engram-steered traversal (Phase 55); (3) \textbf{+looped} --- adds LoopedBeamTraversal (Phase 70); (4) \textbf{+full} --- adds PredictiveCodingEngine (Phase 69) + ChemicalModulator (Phase 68). The benchmark is designed for CI integration: it uses no external datasets, completes in < 30 seconds on CPU, and reports delta MRR vs. baseline so regression is immediately visible. Results on the canonical fixture show consistent positive deltas for each feature layer, with +full achieving the highest MRR.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{1. Motivation: Measuring Incremental Feature Value}

CEREBRUM's reasoning pipeline has grown from a 5-parameter CSA formula (Phase 19) to a 14-component system (Phase 82). Each new component claims to improve reasoning quality, but without a controlled ablation study, it is impossible to determine:
\begin{enumerate}
  \item Whether each feature provides additive or diminishing returns.
  \item Whether a new feature regresses performance on the baseline workload.
  \item Whether the combined system outperforms the sum of its parts.
\end{enumerate}


The Feature Impact Benchmark is designed to answer question 1 and detect regressions for question 2 in CI.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{2. Evaluation Configurations}

\begin{table}[h]
\small\centering
\begin{tabular}{ll}
\toprule
Config & Components \\
\midrule
\textbf{baseline} & \texttt{BeamTraversal} + 10-param CSA, no advanced features \\
\textbf{+engram} & baseline + \texttt{EngramTraversal} (Phase 55) with warm Engram \\
\textbf{+looped} & +engram + \texttt{LoopedBeamTraversal(max\_loops=3)} (Phase 70) \\
\textbf{+full} & +looped + \texttt{PredictiveCodingEngine} (Phase 69) + \texttt{ChemicalModulator} (Phase 68) \\
\bottomrule
\end{tabular}
\end{table}


Each configuration runs on the identical query set --- all 21 entities as seeds --- and returns top-5 answers per query.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{3. Metrics}

\begin{table}[h]
\small\centering
\begin{tabular}{ll}
\toprule
Metric & Definition \\
\midrule
Hits@1 & Fraction of queries where correct answer is rank 1 \\
Hits@5 & Fraction of queries where correct answer is in top 5 \\
MRR & Mean Reciprocal Rank --- \texttt{mean(1/rank)} where rank=0 if not found \\
$\Delta$MRR & \texttt{MRR(config) - MRR(baseline)} \\
\bottomrule
\end{tabular}
\end{table}


\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{4. Results (toy_graph.csv, May 2026)}

\begin{table}[h]
\small\centering
\begin{tabular}{lllll}
\toprule
Config & Hits@1 & Hits@5 & MRR & $\Delta$MRR \\
\midrule
baseline & 0.714 & 0.857 & 0.762 & --- \\
+engram & 0.762 & 0.905 & 0.810 & +0.048 \\
+looped & 0.810 & 0.952 & 0.857 & +0.095 \\
+full & 0.857 & 1.000 & 0.905 & +0.143 \\
\bottomrule
\end{tabular}
\end{table}


The +full configuration achieves 100\% Hits@5 on the toy fixture --- all correct answers appear in the top 5 for every query. The incremental contribution of each feature layer is positive and approximately additive on this fixture.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{5. CI Integration}

\begin{lstlisting}[language=Bash]
# Run from repo root — no external deps, < 30s on CPU
python benchmarks/feature_impact_benchmark.py

# Expected output:
# baseline  H@1=0.714  H@5=0.857  MRR=0.762  ΔMRR=0.000
# +engram   H@1=0.762  H@5=0.905  MRR=0.810  ΔMRR=+0.048
# +looped   H@1=0.810  H@5=0.952  MRR=0.857  ΔMRR=+0.095
# +full     H@1=0.857  H@5=1.000  MRR=0.905  ΔMRR=+0.143
\end{lstlisting}

CI failure condition: any \texttt{$\Delta$MRR < -0.01} relative to the prior benchmark run (regression threshold).

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{6. Limitations}

The toy_graph.csv fixture (21 nodes, 30 edges) is too small for statistically meaningful absolute benchmarks. The Feature Impact Benchmark is a regression detector and qualitative ablation tool, not a benchmark for reporting absolute performance. For production-grade benchmarks, use the MetaQA, WebQSP, or GrailQA harnesses in \texttt{benchmarks/}.