
\section*{Abstract}

We present \textbf{Loop-Provenance Recovery}, an integration between \texttt{AutonomousDiscoveryLoop} (Phase 74) and \texttt{ProvenanceLedger} (Phase 76) that automatically undoes graph materializations when the circuit breaker fires. Prior to Phase 79, a circuit breaker trip paused future materialization but left already-materialized edges from the failed cycle in the graph. These edges may represent low-quality discoveries that degraded graph precision. Loop-Provenance Recovery addresses this by wiring \texttt{LoopConfig.auto\_rollback\_on\_trip=True} to an automatic call to \texttt{ProvenanceLedger.rollback\_cycle()} at trip time. \texttt{CycleRecord.edges\_rolled\_back} records how many edges were removed, providing observability. The combined system achieves \textbf{fail-safe materialization}: if a discovery cycle degrades below quality thresholds, both future materializations pause \textit{and} past materializations are undone --- returning the graph to its pre-cycle state.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{1. Motivation: Incomplete Recovery}

The Phase 74 circuit breaker addresses the \textit{forward} problem: when approval rate drops below \texttt{min\_approval\_rate}, future materialization pauses. However, the cycle that triggered the trip has already materialized some edges. If those edges are low-quality (e.g., confidences just above threshold on marginally-approved findings), they remain in the graph indefinitely.

In a production deployment with continuous discovery, a single degraded cycle can introduce dozens of noisy edges before the circuit breaker fires. Manual cleanup requires identifying the exact edges from that cycle --- a non-trivial task without provenance tracking.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{2. Recovery Protocol}

When the circuit breaker trips:

\begin{lstlisting}
1. circuit_breaker_tripped = True → pause future materializations
2. if auto_rollback_on_trip and provenance_ledger is not None:
       rolled_back = provenance_ledger.rollback_cycle(cycle_number, adapter)
       cycle_record.edges_rolled_back = rolled_back
3. log: "Circuit breaker tripped. Rolled back {rolled_back} edges from cycle {N}."
4. sleep(next_interval)
5. Next cycle: check if approval rate has recovered
\end{lstlisting}

\subsection*{Prerequisite: ProvenanceLedger Attachment}

\begin{lstlisting}[language=Python]
ledger = ProvenanceLedger(max_batches=500)
research_agent.set_provenance_ledger(ledger)
loop = AutonomousDiscoveryLoop(
    agent=research_agent,
    config=LoopConfig(auto_rollback_on_trip=True),
    auto_approver=approver
)
\end{lstlisting}

If \texttt{provenance\_ledger} is \texttt{None}, the auto-rollback silently skips (backward-compatible behavior).

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{3. CycleRecord Fields}

\begin{lstlisting}[language=Python]
@dataclass
class CycleRecord:
    ...
    circuit_breaker_tripped: bool
    edges_rolled_back: int = 0   # 0 if not tripped or rollback not configured
\end{lstlisting}

The Studio v2 cycle history panel (Phase 75) renders \texttt{edges\_rolled\_back} in the cycle table, making rollback events visible to operators without requiring log inspection.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{4. GraphAdapter Prerequisite}

\texttt{rollback\_cycle()} calls \texttt{adapter.remove\_edge(u, v, relation)} for each edge in the cycle's batches. This requires the adapter to implement the Phase 80 \texttt{remove\_edge()} protocol. All built-in adapters support this. Custom adapters must implement it or raise \texttt{NotImplementedError} explicitly.

\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{5. Failure Modes and Mitigations}

\begin{table}[h]
\small\centering
\begin{tabular}{lll}
\toprule
Failure & Behavior & Mitigation \\
\midrule
Partial rollback (adapter exception mid-cycle) & \texttt{rollback\_cycle} catches exception, logs warning, continues with remaining edges & Idempotent: re-running \texttt{rollback\_batch} on completed batches is a no-op \\
ProvenanceLedger evicted cycle batches (LRU cap) & Rollback is incomplete --- only unevicted batches are removed & Tune \texttt{max\_batches} to retain at least \texttt{circuit\_breaker\_window} cycles \\
Adapter does not implement \texttt{remove\_edge} & \texttt{NotImplementedError} propagates, trip still recorded & Implement \texttt{remove\_edge} or disable \texttt{auto\_rollback\_on\_trip} \\
\bottomrule
\end{tabular}
\end{table}


\medskip\noindent\rule{\linewidth}{0.4pt}\medskip

\section*{6. References}

\begin{itemize}
  \item Phase 74 (PAPER_027): AutonomousDiscoveryLoop circuit breaker
  \item Phase 76 (PAPER_029): ProvenanceLedger data model and rollback API
  \item Phase 80 (PAPER_032): GraphAdapter \texttt{remove\_edge} protocol
\end{itemize}
