\section{CEREBRUM: Academic Source Repository (Batch Index)}

This document provides a conso\-lidated list of the 45 found\-ational papers, books, and datasets cited in the CEREBRUM framework development. Each entry is hyperlinked to its official digital record or Open Access PDF.

\begin{itemize}
\item \textbf{hebb1949}: Hebb, D. (1949). The organ\-ization of behavior. \href{https://archive.org/details/organizationofbe00hebb}{View Source}
\item \textbf{vaswani\-2017\-attention}: Vaswani et al. (2017). Attention is all you need. \href{https://arxiv.org/pdf/1706.03762.pdf}{Download PDF}
\item \textbf{bipoo1998}: Bi \& Poo (1998). Synaptic modif\-ications in neurons. \href{https://www.jneurosci.org/content/jneuro/18/24/10464.full.pdf}{Download PDF}
\item \textbf{traag\-2019\-louvain}: Traag et al. (2019). From Louvain to Leiden. \href{https://arxiv.org/pdf/1810.08473.pdf}{Download PDF}
\item \textbf{blondel\-2008\-louvain}: Blondel et al. (2008). Fast unfolding of communities. \href{https://arxiv.org/pdf/0803.0476.pdf}{Download PDF}
\item \textbf{raghavan\-2007\-lpa}: Raghavan et al. (2007). Near linear time community detection. \href{https://arxiv.org/pdf/0709.2931.pdf}{Download PDF}
\item \textbf{page\-1999\-pagerank}: Page et al. (1999). The PageRank citation ranking. \href{http://ilpubs.stanford.edu:8090/422/1/1999-66.pdf}{Download PDF}
\item \textbf{thompson\-1933\-bayesian}: Thompson, W. (1933). On the likelihood of unknown probability. \href{https://www.gwern.net/docs/statistics/bayes/1933-thompson.pdf}{Download PDF}
\item \textbf{russo\-2018\-thompson}: Russo et al. (2018). A tutorial on Thompson sampling. \href{https://web.stanford.edu/~bvr/pubs/TS_Tutorial.pdf}{Download PDF}
\item \textbf{schon\-emann\-1966\--procrustes}: Schonemann, P. (1966). Normalized solution of Procrustes. \href{https://psychometrika.org/index.php/psychometrika/article/download/10.1007/BF02289451/1056}{Download PDF}
\item \textbf{gower\-2004\-procr\-ustes}: Gower \& Dijkst\-erhuis (2004). Procrustes problems. \href{https://openlibrary.org/books/OL8850117M/Procrustes_problems}{View Source}
\item \textbf{ying\-2019\-gnnex\-plainer}: Ying et al. (2019). Gnnexp\-lainer. \href{https://arxiv.org/pdf/1903.03894.pdf}{Download PDF}
\item \textbf{ribeiro\-2016\-lime}: Ribeiro et al. (2016). "Why should I trust you?" LIME. \href{https://arxiv.org/pdf/1602.04938.pdf}{Download PDF}
\item \textbf{bordes\-2013\-transe}: Bordes et al. (2013). Translating embeddings (TransE). \href{https://papers.nips.cc/paper/5071-translating-embeddings-for-modeling-multi-relational-data.pdf}{Download PDF}
\item \textbf{sun\-2019\-rotate}: Sun et al. (2019). RotatE: Relational Rotation. \href{https://arxiv.org/pdf/1902.10197.pdf}{Download PDF}
\item \textbf{lacroix\-2020\-tntco\-mplex}: Lacroix et al. (2020). Temporal Knowledge Base Completion. \href{https://openreview.net/pdf?id=Hye_S0NKvB}{Download PDF}
\item \textbf{finn\-2017\-maml}: Finn et al. (2017). MAML meta-learning. \href{https://arxiv.org/pdf/1703.03400.pdf}{Download PDF}
\item \textbf{horrocks\-2004\-owl}: Horrocks et al. (2004). The OWL Web Ontology Language. \href{https://www.w3.org/TR/owl-features/}{View Source}
\item \textbf{metaqa2017}: Zhang et al. (2018). Variational reasoning (MetaQA). \href{https://arxiv.org/pdf/1709.04071.pdf}{Download PDF}
\item \textbf{webqsp2016}: Yih et al. (2016). Semantic parse labeling (WebQSP). \href{https://aclanthology.org/P16-2033.pdf}{Download PDF}
\item \textbf{hetionet\-2017\-}: Himmelstein et al. (2017). Systematic integration (Hetionet). \href{https://elifesciences.org/articles/26726.pdf}{Download PDF}
\item \textbf{miller2019-expla\-inability}: Miller, T. (2019). Explanation in AI: Social Sciences. \href{https://arxiv.org/pdf/1706.07269.pdf}{Download PDF}
\item \textbf{samek2017-explainable}: Samek et al. (2017). Explainable AI foundations. \href{https://arxiv.org/pdf/1708.08296.pdf}{Download PDF}
\item \textbf{lundberg\-2017\-shap}: Lundberg \& Lee (2017). A unified approach (SHAP). \href{https://papers.nips.cc/paper/7062-a-unified-approach-to-interpreting-model-predictions.pdf}{Download PDF}
\item \textbf{rosvall\-2008\-infomap}: Rosvall et al. (2008). Maps of random walks (Infomap). \href{https://www.pnas.org/content/pnas/105/4/1118.full.pdf}{Download PDF}
\item \textbf{fortunato\-2010\--community}: Fortunato, S. (2010). Community detection in graphs. \href{https://arxiv.org/pdf/0906.0612.pdf}{Download PDF}
\item \textbf{reimers\-2019\-sente\-ncebert}: Reimers et al. (2019). Sentence-BERT. \href{https://arxiv.org/pdf/1908.10084.pdf}{Download PDF}
\item \textbf{velic\-kovic\-2018\-gat}: Velickovic et al. (2018). Graph Attention Networks (GAT). \href{https://arxiv.org/pdf/1710.10903.pdf}{Download PDF}
\item \textbf{trouillon\-2016\-complex}: Trouillon et al. (2016). Complex embeddings (ComplEx). \href{https://arxiv.org/pdf/1606.06357.pdf}{Download PDF}
\item \textbf{nichol\-2018\-reptile}: Nichol et al. (2018). On first-order meta-learning (Reptile). \href{https://arxiv.org/pdf/1803.02999.pdf}{Download PDF}
\item \textbf{jin\-2020\-renet}: Jin et al. (2020). Recurrent event network (RENet). \href{https://arxiv.org/pdf/2004.13524.pdf}{Download PDF}
\item \textbf{qu2020tdgnn}: Qu et al. (2020). Temporal dynamics-aware GNN. \href{https://arxiv.org/pdf/2002.04873.pdf}{Download PDF}
\item \textbf{dong\-2014\-knowl\-edgevault}: Dong et al. (2014). Knowledge Vault. \href{https://research.google/pubs/knowledge-vault-a-web-scale-infrastructure-for-data-fusion.pdf}{Download PDF}
\item \textbf{bertossi\-2011\-database}: Bertossi, L. (2011). Database incon\-sistency constraints. \href{https://www.morganclaypool.com/doi/abs/10.2200/S00305ED1V01Y201010DTM011}{View Source}
\item \textbf{bloom1970}: Bloom, B. (1970). Bloom Filter. \href{http://crystal.uta.edu/~mcguigan/cse6339/papers/Bloom70.pdf}{Download PDF}
\item \textbf{dieke\-lmann\-2010\-memory}: Diekelmann \& Born (2010). The memory function of sleep. \href{https://www.nature.com/articles/nrn2762.pdf}{Download PDF}
\item \textbf{doan\-2012\-princ\-iples}: Doan et al. (2012). Principles of data integration. \href{https://www.sciencedirect.com/book/9780124158146/principles-of-data-integration}{View Source}
\item \textbf{gentry2009}: Gentry, C. (2009). Fully homomorphic encryption. \href{https://crypto.stanford.edu/craig/craig-thesis.pdf}{Download PDF}
\item \textbf{wu2012probase}: Wu et al. (2012). Probase: Probab\-ilistic taxonomy. \href{https://www.cs.uwyo.edu/~ruben/cs5011/probase.pdf}{Download PDF}
\item \textbf{carlson\-2010\-nell}: Carlson et al. (2010). Never-ending language learning (NELL). \href{http://www.cs.cmu.edu/~acarlson/papers/carlson-aaai10.pdf}{Download PDF}
\item \textbf{yang\-2013\-bigclam}: Yang \& Leskovec (2013). Overlapping community (BIGCLAM). \href{https://arxiv.org/pdf/1205.6228.pdf}{Download PDF}
\item \textbf{coscia\-2012\-demon}: Coscia et al. (2012). Local-first overlapping (DEMON). \href{https://arxiv.org/pdf/1206.0629.pdf}{Download PDF}
\item \textbf{lin\-2015\-ttranse}: Lin et al. (2015). Knowledge graph completion (TTransE). \href{https://thunlp.org/~lzy/publications/aaai2015_transr.pdf}{Download PDF}
\item \textbf{sun2017hyte}: Sun et al. (2017). Temporally Aware Embedding (HyTE). \href{https://www.researchgate.net/profile/Shibiao_Sun/publication/320448102_Hyte_Joint_embedding_of_entities_relations_and_time/links/5a9e59d9a6fdcc546ad881f3/Hyte-Joint-embedding-of-entities-relations-and-time.pdf}{Download PDF}
\item \textbf{saxena\-2020\-improve}: Saxena et al. (2020). Improving multi-hop QA (EmbedKGQA). \href{https://arxiv.org/pdf/1909.11855.pdf}{Download PDF}

\end{itemize}